---
title: "This is Tranmere's best National League XI according to PPG"
description: "They fit nicely into a 4-4-2 with a very handy subs' bench"
author: "Pete Brown"
date: "03/03/2023"
draft: true
---

```{r}
#| echo: false
#| message: false

library(tidyverse)
library(knitr)
library(kableExtra)

players <- read_csv("https://raw.githubusercontent.com/petebrown/update-player-stats/main/data/players_df.csv", show_col_types = FALSE) %>%
    rename(player_id = sb_player_id,
           game_id = sb_game_id) %>%
    mutate(player_id = as.numeric(player_id),
           game_id = str_replace_all(game_id, "tpg",""),
           game_id = as.numeric(game_id)) %>%
    select(game_id, game_date, player_id, player_name)

subs <- read_csv("https://raw.githubusercontent.com/petebrown/scrape-events/main/data/subs-and-reds.csv", show_col_types = FALSE)

results <- read_csv("https://raw.githubusercontent.com/petebrown/update-results/main/data/results_df.csv", show_col_types = FALSE) %>%
       select(game_date, opposition, outcome, league_tier)

df <- players %>%
    full_join(subs, by = c("game_id", "player_id")) %>%
    left_join(results, by = "game_date") %>%
    filter(!is.na(league_tier)) %>%
    mutate(points = case_when(outcome == "W" ~ 3,
                              outcome == "D" ~ 1,
                              TRUE ~ 0),
           role = case_when(is.na(min_on) ~ "starter",
                            TRUE ~ "sub")) %>%
    mutate(mins_played = case_when(role == "starter" & is.na(min_off) ~ 90,
                                   role == "starter" & !is.na(min_off) ~ min_off,
                                   role == "starter" & !is.na(min_so) ~ min_so,
                                   role == "sub" & is.na(min_so) ~ 90 - min_on,
                                   role == "sub" & !is.na(min_so) ~ min_so - min_on)) %>%
    select(player_name, game_date, opposition, league_tier, outcome, points, role, min_on, min_off, min_so, mins_played)

nl <- df %>%
    filter(league_tier == 5)

av_pts <- nl %>%
    group_by(player_name) %>%
    filter(role == "starter") %>%
    summarise(starts = sum(role == "starter"),
              wins = sum(outcome == "W"),
              draws = sum(outcome == "D"),
              losses = sum(outcome == "L"),
              win_pc = wins / starts,
              av_pts = sum(points) / starts) %>%
    filter(starts >= 10) %>%
    ungroup() %>%
    mutate(rank = min_rank(desc(av_pts)),
           win_pc = str_glue("{round(win_pc * 100, 0)}%")) %>%
    arrange(desc(av_pts), desc(win_pc)) %>%
    mutate(av_pts = round(av_pts, 2),
           ssn_pts = av_pts * 46)
```

### Who makes Rovers' National League All-Star XI?
```{r}
#| echo: false
#| message: false

starters <- av_pts %>%
    select(rank, player_name, starts, av_pts, ssn_pts) %>%
    rename(Rank = rank,
           Player = player_name,
           Starts = starts,
           PPG = av_pts) %>%
    filter(Rank <= 10 | Player == "Scott Davies")

starters %>%
    select(-ssn_pts) %>%
    knitr::kable() %>%
    kable_styling(bootstrap_options = "striped", "responsive", full_width = F, position = "center")
```

Davies

Vaughan - Ihiekwe - Buxton - Clarke

Dawson - Hughes - Maynard - Tollitt

Jennings

Cook

### Who's on the bench?

The next five make a very handy subs' bench, particularly if we're chasing a game.

```{r}
#| echo: false
#| message: false

subs <- av_pts %>%
    select(rank, player_name, starts, av_pts) %>%
    rename(Rank = rank,
           Player = player_name,
           Starts = starts,
           PPG = av_pts) %>%
    filter(Rank >= 11 & Rank <= 16 & Player != "Scott Davies")
    
subs %>%
    knitr::kable() %>%
    kable_styling(bootstrap_options = "striped", "responsive", full_width = F, position = "center")
```

### Where would this team have finished in the table?

Over a 46 game season, this XI would have accumulated an average of `r round(mean(starters$ssn_pts), 2)` points.

In 2017/18, the season Rovers finally secured promotion via the play-offs, 93 points would have been enough to pip Macclesfield Town to the championship by one point.

| Season  | Position for 93 points | Actual finish   |
|---------|------------------------|-----------------|
| 2015/16 | 2nd                    | 6th (78 points) |
| 2016/17 | 3rd                    | 2nd (95 points) |
| 2017/18 | 1st                    | 2nd (82 points) |

### Here's the XI that played the most minutes in the National League
```{r}
#| echo: false
#| message: false

most_mins <- df %>%
    filter(league_tier == 5) %>%
    group_by(player_name) %>%
    summarise(total_mins = sum(mins_played)) %>%
    arrange(desc(total_mins)) %>%
    ungroup() %>%
    mutate(rank = min_rank(desc(total_mins))) %>%
    filter(rank <= 11)

most_mins %>%
    select(rank, player_name, total_mins) %>%
    rename("Player Name" = player_name,
           "Total mins" = total_mins,
           "Rank" = rank) %>%
    knitr::kable(align = "rlr", format.args = list(big.mark = ",")) %>%
    kable_styling(bootstrap_options = "striped", "responsive", full_width = F, position = "center")
```

Davies

Sutton - McNulty - Ihekwe - Ridehalgh

Hughes

Harris - Maynard

Jennings

Norwood - Cook

Ihekwe, Hughes, Maynard, Jennings and Cook